---
title: ""
author: Miguel Ferris
date: 20-06-2019
output:
  html_document:
    code_folding: hide
  pdf_document:
    toc: no
---

# MY PROJECTS {.tabset}
##  Introduction{.tabset}
<br>
This document shows personal projects done in order to practice and gain knowledge about different technologies and problems.
<br>


1. <font size="4">**Mercado Spot:**</font>
   +  **Description:**<br>
  In this project the Spanish electric market is analyzed, to try to determine what parameters affect the price of electricity and if this can be predicted.
  + **Knowledge and Tecnoligies:**<br>
    R programing, MachineLearning, Data treatment, Electric market
    <br>
2. <font size="4">**Dashboard Project:**</font>
  + **Description:**<br>
    This project presents a computer tool based on a web application for the visualization and control of indicators and important parameters in the welding process. 
  + **Knowledge and Tecnoligies:**<br>
    R programing, Shiny Package, Visualization
<br>
3. <font size="4">**Vehicle Routing Problem - VRP:**</font>
  + **Description:**<br>
    The problem to which this program responds is the picking up of several people in their locality of residence in the morning and return them at the end of the day. To do this, they have a fleet with different types of vehicles, which should be combined in the best possible way trying to minimize the inverted times, kilometers traveled, etc.
  + **Knowledge and Tecnoligies:**<br>
    Excel, VBA programing,Combinatorial optimization

    
##  Mercado Spot{.tabset}
**Data Science Framework**<br>
1. Define the Problem <br>
2. Gather the Data: <br>
3. Prepare Data for Consumption<br>
4. Perform Exploratory Analysis<br>
5. Model Data<br>
6. Validate and Implement Data Model<br>
7. Optimize and Strategize<br>
<br>

<font size="4">**1. Define the Problem**</font><br>
Analyze and find the parameters that are more influential in the prize of electricity. For later create a model that tries to predict the price.

<font size="4">**2. Gather the Data**</font><br>
The data is gathered from ESIOS, a public platform where anyone, for free, can view and download information related to the electrical system.
```{r message=FALSE, warning = FALSE}
#ESIOS
ESIOS<-read.csv('/Users/miguelferrisllandete/Desktop/data_analisis/Sistema electrico/m3/libro4.csv',sep=";",stringsAsFactors = FALSE)
library(tidyr)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(plotly)
library(lubridate)
library(Metrics)
library("rpart")
library("rpart.plot")

```


<font size="4">**3. Prepare Data for Consumption**</font>
<br>
3.1. Change the format of the date column and create two extra columns mes(month) and dia(day) 
```{r message=FALSE, warning = FALSE}
# Convert Value to numeric
ESIOS$value<-as.numeric(gsub(",","",ESIOS$value))
# Data format transformations
ESIOS<-spread(ESIOS,key=name,value=value)
ESIOS$datetime<-(gsub("T"," ",ESIOS$datetime))
ESIOS$datetime<-gsub("f01:00","",ESIOS$datetime)
ESIOS$datetime<-as.POSIXct(ESIOS$datetime,"%Y-%m-%d %H:%M:%OS",tz="Europe/Madrid")
ESIOS<-ESIOS%>%mutate(mes=as.numeric(format(ESIOS$datetime,'%m')))
ESIOS<-ESIOS%>%mutate(Hora=as.numeric(format(ESIOS$datetime,'%H')))
ESIOS<-ESIOS[-1,]

```


3.2. Featuring Engineering: 

+ **Precio_dia_anterior:** Gives The price of the day before.<br>
+ **Precio_semana_anterior:** Gives The price of the week before.<br>
+ **Precio_año_anterior:** Gives The price of the year before.<br>
+ **Dia:** column that says if the day is a holiday or not.<br>
+ **Estacion:** Winter/Summer.<br>
+ **Periodo:** Classify the hours of a non holiday day into 3 periods and Classify the hours of a holiday day into 2 periods.<br>
```{r message=FALSE, warning = FALSE}


#Create column that gives The price of the day before
x<-ifelse(as.numeric(row.names(ESIOS))-25<=0,as.numeric("2"),as.numeric(row.names(ESIOS))-25)
ESIOS$Precio_dia_anterior<-ESIOS[x,"Precio mercado SPOT Diario"]

#Create column that gives The price of the week before
x<-ifelse(as.numeric(row.names(ESIOS))-169<=0,as.numeric("2"),as.numeric(row.names(ESIOS))-169)
ESIOS$Precio_semana_anterior<-ESIOS[x,"Precio mercado SPOT Diario"]

#Create column that gives The price of the year before
x<-ifelse(as.numeric(row.names(ESIOS))-8761<=0,as.numeric("2"),as.numeric(row.names(ESIOS))-8761)
ESIOS$Precio_año_anterior<-ESIOS[x,"Precio mercado SPOT Diario"]

#Create column that says if the day is a holiday or not
ESIOS<-ESIOS%>%mutate(dia=(format(ESIOS$datetime,'%a')))
ESIOS[!ESIOS$dia%in%c("dom","sáb"),"dia"]<-"No_Festivo"
ESIOS[ESIOS$dia%in%c("dom","sáb"),"dia"]<-"Festivo"

#Winter/Summer
x<-unlist(lapply(ESIOS$mes,function(x) if(x%in%c(11,12,1,2,3)){"Invierno"}else{"Verano"}))
ESIOS$Estacion<-x
ESIOS$Estacion<-as.factor(ESIOS$Estacion)

#Classify the hours of a non holiday day into 3 periods
y<-ifelse(ESIOS$Hora%in%c(1,2,3,4,5,6,7,8),"P3",ifelse((ESIOS$Hora%in%c(9,10,11,12,13,14,15,16,17,24)&ESIOS$Estacion=="Invierno")|(ESIOS$Hora%in%c(9,10,17,18,19,20,21,22,23,24)&ESIOS$Estacion=="Verano"),"P2","P1"))
ESIOS$Periodo<-y

#Classify the hours of a holiday day into 2 periods
ESIOS[ESIOS$dia=="Festivo"&ESIOS$Hora%in%c(1:18),
               "Periodo"]<-"P3"
ESIOS[ESIOS$dia=="Festivo"&ESIOS$Hora%in%c(19:24),
               "Periodo"]<-"P2"

# Get data only from year=2018
ESIOS2018<-ESIOS[ESIOS$datetime>="2018-01-01" & ESIOS$datetime<="2018-12-31",]

names(ESIOS2018)<-gsub(" ","_",names(ESIOS2018))
ESIOS2018$dia<-as.factor(ESIOS2018$dia)
ESIOS2018$Periodo<-as.factor(ESIOS2018$Periodo)

knitr::kable(
  ESIOS[1:6,], caption = 'A subset of the Dataset.'
)
```

<font size="4">**4. Exploratory analysis**</font><br>


#### 4.1 Target Variable (Precio_mercado_SPOT_Diario) 

```{r message=FALSE, warning = FALSE}
summary(ESIOS2018$Precio_mercado_SPOT_Diario)
```


```{r message=FALSE, warning = FALSE}
ggplot(data.frame(x = ESIOS2018$Precio_mercado_SPOT_Diario), aes(x = x)) +
    geom_histogram(binwidth = 3) +
  labs(x = "price", y = "Frequency")+theme_minimal()

```

+ Deviate from the normal distribution.

+ Have appreciable negative skewness.

<br>

#### 4.2 Numeric Variables {.tabset .tabset-fade}

<br>
<br>
To start exploring the data with descriptive and graphical statistics. We are going to determining the correlation between each numeric variable and the target variable .
<br>
Scatter charts can help us to understand how the prize of electricity dependeds with the different features.
<br>
<br>

##### Forecast Demand 
<br>
Pearson correlation:
```{r message=FALSE, warning = FALSE}
cor(ESIOS2018$Forecasted_demand, ESIOS2018$Precio_mercado_SPOT_Diario)
```

Scatterplots:

```{r message=FALSE, warning = FALSE}
ggplot(ESIOS2018,aes(x=Forecasted_demand,y=Precio_mercado_SPOT_Diario))+
  geom_point(position = "jitter", alpha = 0.3)+geom_smooth(method = lm)+theme_minimal()
```

+ 0.41 A moderate uphill (positive) relationship

##### Last day Price
<br>
Pearson correlation:
```{r, warning = FALSE}
cor(ESIOS2018$Precio_dia_anterior, ESIOS2018$Precio_mercado_SPOT_Diario)
```
Scatterplots:
```{r, warning = FALSE}
ggplot(ESIOS2018,aes(x=Precio_dia_anterior,y=Precio_mercado_SPOT_Diario))+
  geom_point(position = "jitter", alpha = 0.3)+geom_smooth(method = lm)+theme_minimal()
```

+ 0.8 A strong uphill (positive) linear relationship

##### Last week Price
<br>
Pearson correlation:
```{r message=FALSE, warning = FALSE}
cor(ESIOS2018$Precio_semana_anterior, ESIOS2018$Precio_mercado_SPOT_Diario)
```
Scatterplots:
```{r, warning = FALSE}
ggplot(ESIOS2018,aes(x=Precio_semana_anterior,y=Precio_mercado_SPOT_Diario))+
  geom_point(position = "jitter", alpha = 0.3)+geom_smooth(method = lm)+theme_minimal()
```

+ 0.67 A strong uphill (positive) linear relationship

##### Last year Price
<br>
Pearson correlation:
```{r message=FALSE, warning = FALSE}
cor(ESIOS2018$Precio_año_anterior, ESIOS2018$Precio_mercado_SPOT_Diario)
```
Scatterplots:
```{r, warning = FALSE}
ggplot(ESIOS2018,aes(x=Precio_año_anterior,y=Precio_mercado_SPOT_Diario))+
  geom_point(position = "jitter", alpha = 0.3)+geom_smooth(method = lm)+theme_minimal()
```

+ 0.29 A weak uphill (positive) linear relationship


##### Forecast wind power generation
<br>
Pearson correlation:
```{r message=FALSE, warning = FALSE}
cor(ESIOS2018$Peninsular_wind_power_generation_forecast, ESIOS2018$Precio_mercado_SPOT_Diario)
```
Scatterplots:
```{r, warning = FALSE}
ggplot(ESIOS2018,aes(x=Peninsular_wind_power_generation_forecast,y=Precio_mercado_SPOT_Diario))+
  geom_point(position = "jitter", alpha = 0.3)+geom_smooth(method = lm)+theme_minimal()
```

+ 0.41 A moderate uphill (positive) relationship

##### Forecast solar generation
<br>
Pearson correlation:
```{r message=FALSE, warning = FALSE}
cor(ESIOS2018$Solar_PV_generation_forecast, ESIOS2018$Precio_mercado_SPOT_Diario)
```
Scatterplots:
```{r, warning = FALSE}
ggplot(ESIOS2018,aes(x=Solar_PV_generation_forecast,y=Precio_mercado_SPOT_Diario))+
  geom_point(position = "jitter", alpha = 0.3)+geom_smooth(method = lm)+theme_minimal()
```

+ 0.1 A realy weak uphill (positive) linear relationship


##### Forecast thermal solar generation
<br>
Pearson correlation:
```{r message=FALSE, warning = FALSE}
cor(ESIOS2018$Solar_thermal_forecast, ESIOS2018$Precio_mercado_SPOT_Diario)
```
Scatterplots:
```{r, warning = FALSE}
ggplot(ESIOS2018,aes(x=Solar_thermal_forecast,y=Precio_mercado_SPOT_Diario))+
  geom_point(position = "jitter", alpha = 0.3)+geom_smooth(method = lm)+theme_minimal()
```

+ 0.175 A realy weak uphill (positive) linear relationship


#### 4.3 Categorical Variables {.tabset .tabset-fade}

##### Weekend / No-Weekend
<br>
Box plot:
```{r message=FALSE, warning = FALSE}
ggplot(ESIOS2018,aes(y=Precio_mercado_SPOT_Diario,x=as.factor(dia),fill=as.factor(dia)))+
  geom_boxplot()+theme_minimal()
  
```

```{r message=FALSE, warning = FALSE}
anova <- aov(ESIOS2018$Precio_mercado_SPOT_Diario ~ ESIOS2018$dia)
summary(anova)
```

##### Winter / Summer
<br>
Box plot:
```{r message=FALSE, warning = FALSE}
ggplot(ESIOS2018,aes(y=Precio_mercado_SPOT_Diario,x=as.factor(Estacion),fill=as.factor(Estacion)))+
  geom_boxplot()+theme_minimal()
```

```{r message=FALSE, warning = FALSE}
anova <- aov(ESIOS2018$Precio_mercado_SPOT_Diario ~ ESIOS2018$dia)
summary(anova)
```

##### Periods of hours
<br>
Box plot:
```{r message=FALSE, warning = FALSE}
ggplot(ESIOS2018,aes(y=Precio_mercado_SPOT_Diario,x=as.factor(Periodo),fill=as.factor(Periodo)))+
  geom_boxplot()+theme_minimal()
```

##### Month
<br>
Box plot:
```{r message=FALSE, warning = FALSE}
ggplot(ESIOS2018,aes(y=Precio_mercado_SPOT_Diario,x=as.factor(mes),fill=as.factor(mes)))+
  geom_boxplot()+theme_minimal()

```

```{r message=FALSE, warning = FALSE}
anova <- aov(ESIOS2018$Precio_mercado_SPOT_Diario ~ ESIOS2018$dia)
summary(anova)
```


#### 5. Model Data {.tabset .tabset-fade}

##### TREE
<br>

To start looking into creating a good model, first, we tried with a one tree model.

```{r message=FALSE, warning = FALSE}

pv_model<-ESIOS2018

set.seed(123)
assignment <- sample(1:3, size = nrow(pv_model), prob = c(80,10,10), replace = TRUE)

# Create a train, validation and tests from the original data frame 
pv_model_train <- pv_model[assignment == 1, ]    # subset grade to training indices only
pv_model_valid <- pv_model[assignment == 2, ]  # subset grade to validation indices only
pv_model_test <- pv_model[assignment == 3, ]   # subset grade to test indices only

pv_model_model <- rpart(formula = Precio_mercado_SPOT_Diario ~ ., 
                        data = pv_model_train, 
                        method = "anova")


#Generate predictions on a test set

pred <- predict(object = pv_model_model,   # model object 
                newdata = pv_model_test)  # test dataset
```

Plot the tree model
```{r message=FALSE, warning = FALSE}
rpart.plot(x = pv_model_model, yesno = 2, type = 0, extra = 0)
```

See results and Compute the RMSE
```{r message=FALSE, warning = FALSE}
rmse<-rmse(actual = pv_model_test$Precio_mercado_SPOT_Diario, 
           predicted =pred)
rmse
```

##### RANDOM FOREST

```{r message=FALSE, warning = FALSE}
ESIOS2018<-ESIOS2018%>%mutate(mes=as.factor(mes),Periodo=as.factor(Periodo),dia=as.factor(dia),Estacion=as.factor(Estacion))
Valores_importantes<-ESIOS2018[complete.cases(ESIOS2018),]

library(randomForest)
set.seed(123)
pv_modelR<-Valores_importantes%>%mutate(datetime=as.numeric(datetime))
assignment <- sample(1:2, size = nrow(pv_modelR), prob = c(90,10), replace = TRUE)

# Create a train, validation and tests from the original data frame 
pv_modelR_train <- pv_modelR[assignment == 1, ]    # subset grade to training indices only
#pv_modelR_valid <- pv_modelR[assignment == 2, ]  subset grade to validation indices only
pv_modelR_test <- pv_modelR[assignment == 2, ]   # subset grade to test indices only

# Train a Random Forest
# for reproducibility
precio_volumen2018_Rmodel <- randomForest(formula = Precio_mercado_SPOT_Diario~.,
                                          data = pv_modelR_train,
                                          mtry=4,
                                          nodesize=3,
                                          sampsize=6116,
                                          ntree=400
                                          )

# Print the model output                             
#print(precio_volumen2018_Rmodel)
plot(precio_volumen2018_Rmodel)
#precio_volumen2018_Rmodel$mse
# Generate predicted classes using the model object
precio_volumen2018_Rmodel_prediction <- predict(object = precio_volumen2018_Rmodel,    
                                                newdata = pv_modelR_test) 
# return classification labels
rmse5<-rmse(actual = pv_modelR_test$Precio_mercado_SPOT_Diario, 
            predicted = precio_volumen2018_Rmodel_prediction)
mse5<-mse(actual = pv_modelR_test$Precio_mercado_SPOT_Diario, 
          predicted = precio_volumen2018_Rmodel_prediction)
mae5<-mae(actual = pv_modelR_test$Precio_mercado_SPOT_Diario, 
          predicted = precio_volumen2018_Rmodel_prediction)
# Print the predicted classes
#print(precio_volumen2018_Rmodel_prediction)
varImpPlot(precio_volumen2018_Rmodel)

```


```{r message=FALSE, warning = FALSE}
pv_modelR_test$prediccion<-precio_volumen2018_Rmodel_prediction
real<-pv_modelR_test%>%select(datetime,Precio_mercado_SPOT_Diario)%>%mutate(Ti="R")
prediccion<-pv_modelR_test%>%select(datetime,prediccion)%>%mutate(Ti="P")
names(prediccion)<-names(real)
comparar<-rbind(real,prediccion)

head(pv_modelR_test)
p<-ggplot(comparar,aes(x=datetime,y=Precio_mercado_SPOT_Diario,color=Ti))+
  geom_line()
ggplotly(p,width=1280, height=480)
```



###

<br>
<font size="4">**6. Validate and Implement Data Model**</font><br>
<font size="4">**7. Optimize and Strategize**</font><br>


## Dashboard Project {.tabset}
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
<br>
This video presents a computer tool based on a web application for the visualization and control of indicators and important parameters in the welding process, programmed using the R language over test data. 
<br>
<br>
A brief introduction of what we can see in the video for each page of the application is:
<br>
<br>
**Overview:** This page serves as a cover page and access page, as well as control of some general behavioral parameters.<br>
**Main inconformations:** It contains information about some fundamental parameters for the wellding process. This tab is interactive and allows increasing the level of granularity of the data.<br>
**Auditorium:** The idea behind this is to allow control of the parameters within an interval. For this, the most relevant parameters have been selected, as well as those with standard values. In case these parameters are far from ideal, send automatic emails to those responsible.<br>
**Auditorium configuration:** To modify, delete or add new parameters to the audit control.<br>
**Change Protocol:** It allows to have historical control of all the changes in the parameters.<br>
**Loginformation:** A record is kept of all the points where the user is able to interact with the application. To have a control of possible errors.<br>
<br>
<iframe width="747" height="420" src="https://www.youtube.com/embed/Fk546hIty8k" frameborder="0" allow="accelerometer; autoplay; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

##  Vehicle Routing Problem - VRP{.tabset}
<br>
<br>
The problem that the program of the video responds to, is the picking up of several people in their locality of residence in the morning and return them at the end of the day. To do this, they have a fleet with different types of vehicles, which should be combined in the best possible way trying to minimize the inverted times, kilometers traveled, etc